Dynamic strategies for yaw and induction control of wind farms based on large-eddy simulation and optimization

### 2018-01-02 reply_to_reviewer1.tex

\documentclass[]{article}

\usepackage{color}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage[symbol]{footmisc}
\newcommand{\red}[1]{\textbf{\color{red} #1}}

\usepackage[margin=1.5in]{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0.5\baselineskip}

%opening
\title{Response to Reviewer 1}
\author{}
\date{}

\begin{document}

\maketitle

We are pleased with the positive assessment of our work by the reviewer and thank him/her for thoroughly reading our manuscript and for providing constructive input. The reviewer justly gives suggestions for interesting additional optimal control cases. Although computational costs prohibit us from running these cases in an acceptable timeframe during the revision process, they are subject of current research and will be included in future work. The responses to the reviewer's comments and adaptations in the revised manuscript are discussed below.

\hrulefill

\paragraph{Reviewer} \textit{The present article investigates the improvement on wind-farm power by performing different wind turbine optimization strategies (yaw and induction) modeled using large-eddy simulations (LESs) combined with an adjoint-based optimization method. Two test cases are considered, one characterized by uniform incoming flow conditions (turbulence void), and another where a precursor LES is used to seed inflow turbulence in the domain containing the 4 x 4 wind turbines layout. The authors present new results about combining axial induction and dynamic yaw control. Although for the relatively small wind farm under consideration, the yaw control over-performs induction control, the authors present interesting analysis on the potential of combining the two methods. The manuscript is overall well written and organized, and the majority of the results are correctly interpreted and clearly articulated. Therefore, I recommend publication after minor 	revisions are made. Here below a list a series of concerns/clarifications that need to be addressed prior to its publication.}

\section*{Comments}
\hrulefill

\paragraph{1. Reviewer} \textit{The authors use an actuator disk model that does not account for rotational effects (line 158). However, in lines 362-364 the authors state that their LES model output exhibits curling of the wakes as a result of the control strategy. What is the impact of the nonrotating actuator disk approach on wake curling and turbine control? It would be interesting to repeat on of the cases, in particular I3Y, including the tangential forces on the turbine, and report on the differences in the dynamics as well as power extraction and wind farm efficiency found.}

\paragraph{Response}

\hrulefill

\paragraph{2. Reviewer} \textit{Figure 12a displays a strong power recovery for rows 3 and 4 for the cases where yaw control is applied (this effect is less evident for the induction control cases), which results in a wind farm efficiency increase of ~20\% with respect to the reference case. I wonder how much of that improvement is due to the reduced number of wind turbine rows (only 4), and the ability to quickly adjust to a much higher power extraction level as the end of the wind farm is approached. It would be desirable to perform at least one more simulation (I3Y) with an extended domain and wind farm, doubling the number of turbine rows from 4 to 8. Such case will reveal the trend of the increase in power extraction and wind farm efficiency as the size of the wind farm is enlarged, and will help extrapolate to what the gain would be for a more realistic wind farm size. In addition, the authors could consider another case where the use of periodic boundary conditions in the streamwise direction, as an asymptotic limit for an “infinite” wind farm, allowing to keep the size of the simulation (although not informative of row dependency within the wind farm). These two additional cases will further increase the robustness of the analysis and conclusions regarding the combined induction/yaw control approach proposed, and the finite wind farm effects.}

\paragraph{Response}

\hrulefill

\paragraph{3. Reviewer} \textit{Line 226-230. How is that time related to an eddy and advection scales? What are the implications going to real large-scale forcing variability? Please comment on that.}

\paragraph{Response}

\hrulefill

\paragraph{4. Reviewer} \textit{Line 291-292. It would be beneficial to include the turbulent case also, making it a 4-panel figure. That case is more representative of an ABL flow and therefore provides the reader a better sense of the optimization convergence in a more realistic scenario.}

\paragraph{Response}

\hrulefill

\paragraph{5. Reviewer} \textit{Line 382-385. What about the peak that shows up in all case for St < 0.1 (~0.05)? Please explain.}

\paragraph{Response}

\hrulefill

\end{document}

### 2018-01-02 reply_to_reviewer2.tex

\documentclass[]{article}

\usepackage{color}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage[symbol]{footmisc}
\newcommand{\red}[1]{\textbf{\color{red} #1}}

\usepackage[margin=1.5in]{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0.5\baselineskip}

%opening
\title{Response to Reviewer 2}
\author{}
\date{}

\begin{document}

\maketitle

We are pleased with the positive assessment of our work by the reviewer and thank him/her for thoroughly reading our manuscript and for providing constructive input. The responses to the reviewer's comments and adaptations in the revised manuscript are discussed below.

\hrulefill

\paragraph{Reviewer} \textit{The paper presents very interesting research related to control of wind farms using yaw and induction control mechanism, as well as combination of both. Using large-eddy simulations (LES) combined with adjoint-based optimization the authors showed significant improvement in power generation for 4x4 wind farm layout. Both, uniform and turbulent inflow conditions were considered with 8 different control strategies for each case. The paper is well written and suggested for publication after addressing the following comments:}

\section*{Comments}
\hrulefill

\paragraph{1. Reviewer} \textit{The main novelty in current paper compared to Ref. [14] is the addition of the yawing. This should be clearly stated in the Introduction part, e.g. in the last paragraph (lines 107-118)}

\paragraph{Response}

\hrulefill

\paragraph{2. Reviewer} \textit{Line 137: "...in an accurate model with full state...", references should be provided here with previous studies using this model to show its accuracy. Authors already cited them in Introduction but it would be worth to repeat some of them here.}

\paragraph{Response}

\hrulefill

\paragraph{3. Reviewer} \textit{Line 184: L-BFGS-B. It might be good to spell it out for clarity.}

\paragraph{Response}

\hrulefill

\paragraph{4. Reviewer} \textit{Line 202-203 should be placed before Eq. 13.}

\paragraph{Response}

\hrulefill

\paragraph{5. Reviewer} \textit{Line 211: corresponding notation for dimensions should be provided, or reference to Table 1 should be given right after 4.8x2.4x1}

\paragraph{Response}

\hrulefill

\paragraph{6. Reviewer} \textit{Line 217: Define Zh}

\paragraph{Response}

\hrulefill

\paragraph{7. Reviewer} \textit{Line 215-218: Figure should be provided with 3D domain, to clarify the vertical placement of the turbine.}

\paragraph{Response}

\hrulefill

\paragraph{8. Reviewer} \textit{It might be good to provide brief comments on sensitivity of the solution to control time or explain the choice of current control time. If it was done previously, reference would be enough.}

\paragraph{Response}

\hrulefill

\paragraph{9. Reviewer} \textit{It looks like no towers were included in current simulation. The authors should clearly state it in Introduction and provide comments on importance/non-importance of tower inclusion in Summary part. It could be provided after the authors provided comments on the effect of stratification.}

\paragraph{Response}

\hrulefill

\paragraph{10. Reviewer} \textit{In Table 1: the expression for precursor pressure gradient is inconsistent with the one defined in line 220, i.e. rho is missing.}

\paragraph{Response}

\hrulefill

\paragraph{11. Reviewer} \textit{Line 335: The authors clearly stated that the optimal control is heterogeneous, however no comments on why it is heterogeneous were provided. It was stated that turbine can pick either wake meandering type of motion or wake redirection, however what will determine this decision?}

\paragraph{Response}

\hrulefill

\paragraph{12. Reviewer} \textit{Figure 5: The blue color for C4 and green color for C2 very hard to see, especially on I3Y case.}

\paragraph{Response}

\hrulefill

\paragraph{13. Reviewer} \textit{Figure 6 shows that for uniform inflow in a reference case the turbine in a second row will be producing almost 90\% less power then upstream. As a reader I would be interested to see previous studies on this. Adding corresponding references would be enough.}

\paragraph{Response}

\hrulefill

\paragraph{14. Reviewer} \textit{Figure 7: Add tick labels to X-axis, i.e. to Time. (Same for Fig. 13)}

\paragraph{Response}

\hrulefill

\paragraph{15. Reviewer} \textit{Line 424: "...steady uniform inflow cases", would it be better to use word "constant uniform inflow cases", to avoid any confusion with steady-state analysis.}

\paragraph{Response}

\hrulefill

\paragraph{16. Reviewer} \textit{Line 596: the "and" is missing in "... yaw angle equation \_\_\_ the adjoint gradient"}

\paragraph{Response}

\hrulefill

\paragraph{17. Reviewer} \textit{In Eq. A13 $rho_i$, $gamma_i$ and $chi_i$ should be defined. Currently they are defined in A.4.}

\paragraph{Response}

\hrulefill

\end{document}

### 2018-01-02 reply_to_reviewer3.tex

\documentclass[]{article}

\usepackage{color}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage[symbol]{footmisc}
\newcommand{\red}[1]{\textbf{\color{red} #1}}

\usepackage[margin=1.5in]{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0.5\baselineskip}

%opening
\title{Response to Reviewer 3}
\author{}
\date{}

\begin{document}

\maketitle

We are pleased with the positive assessment of our work by the reviewer and thank him/her for providing constructive input. The responses to the reviewer's comments and adaptations in the revised manuscript are discussed below.

\hrulefill

\paragraph{Reviewer} \textit{This paper presents an interesting study on yaw and induction control of wind farms. The analysis is reliable, and the paper is well written.}

\section*{Comments}
\hrulefill

\paragraph{1. Reviewer} \textit{Page 11, line 325, “additional control cases Ystat and Ymndr…” Although explanations of Ystat and Ymndr can be found later, e.g. on page 20, line 521. It is suggested that these control cases are explained upon their first appearance. In the caption of Figure 6, the case names should be noted as well.}

\paragraph{Response}

\hrulefill

\paragraph{2. Reviewer} \textit{Some figures are difficult to inspect closely. For example, Figures 7, 11, and 13 contain color lines, and the sizes are too small. The authors can present the results in a better manner.}

\paragraph{Response}

\hrulefill

\paragraph{3. Reviewer} \textit{In Figures 9 and 15, the unit of PSD should be added.}

\paragraph{Response}

\hrulefill

\paragraph{4. Reviewer} \textit{Table 2 gives a summary of the effect different control strategies. Why does the meandering yaw case have reduced power gains in the turbulent condition? How do the authors judge the uncertainties in their analysis? For example, by extending the optimization time window or by refining the cell size, will the conclusion be different?}

\paragraph{Response}

\hrulefill

\end{document}

### 2018-01-02 reply_to_reviewer2.tex

\maketitle

We are pleased with the positive assessment of our work by the reviewer and thank him/her for thoroughly reading our manuscript and for providing constructive input. The responses to the reviewer's comments and adaptations in the revised manuscript are discussed below.

\hrulefill

\paragraph{Reviewer} \textit{The paper presents very interesting research related to control of wind farms using yaw and induction control mechanism, as well as combination of both. Using large-eddy simulations (LES) combined with adjoint-based optimization the authors showed significant improvement in power generation for 4 $\times$ 4 wind farm layout. Both, uniform and turbulent inflow conditions were considered with 8 different control strategies for each case. The paper is well written and suggested for publication after addressing the following comments:}

@@

\section*{Comments}
\hrulefill

\paragraph{1. Reviewer} \textit{The main novelty in current paper compared to Ref. [14] is the addition of the yawing. This should be clearly stated in the Introduction part, e.g. in the last paragraph (lines 107-118)}

\paragraph{Response} This is correct. We have made this more explicit in the revised manuscript in lines 107 -- 118 as follows:

``
In the current manuscript, we aim to fill some of the aforementioned gaps in wind-farm control research \textbf{using the dynamic control approach based on large-eddy simulation and optimization as introduced by Goit \& Meyers [11]. The main novelty of the current work is the inclusion of yaw control, in addition to the induction control considered in previous studies [11, 12, 14].} Based on a demonstration $\dots$
''

### 2018-01-02 MuntersMeyers_DynamicStrategiesControl_revision_WM01.tex
R2C1

@@

In the current manuscript, we aim to fill some of the aforementioned gaps in wind-farm control research. We extend the dynamic optimal induction control framework SP--Wind \cite{goit2015optimal, munters2017optimal} to incorporate dynamic yaw control. Based on a demonstration $4 \times 4$ wind-farm control case, we illustrate the benefits of unsteady yaw control, and uncover the potential of combining yaw control with axial induction control. To the best knowledge of the authors, the current manuscript presents a first attempt at quantifying the potential of combining the latter control strategies in a dynamic manner. 
The manuscript is structured as follows. Firstly, Section \ref{sec:meth} introduces the optimal control problem and the numerical methods used to solve it. Next, Section~\ref{sec:opt_yaw_setup} describes the simulation cases, and details setup parameters. Thereafter, Section~\ref{sec:opt_yaw_results} presents optimization results for both uniform and turbulent inflow, and identifies distinct yaw control regimes resulting in increased power extraction. Finally, Section~\ref{sec:opt_yaw_concl} summarizes the main findings of this manuscript and formulates suggestions for future research.

>>>>

\newcommand{\revision}[1]{{\color{red} #1}}

@@

\revision{In the current manuscript, we aim to fill some of the aforementioned gaps in wind-farm control research using the dynamic control approach based on large-eddy simulation and optimization as introduced by Goit \& Meyers \cite{goit2015optimal}. The main novelty of the current work is the inclusion of yaw control, in addition to the induction control considered in previous studies \cite{goit2015optimal, goit2016optimal, munters2017optimal}.} 
Based on a demonstration $4 \times 4$ wind-farm control case, we illustrate the benefits of unsteady yaw control, and uncover the potential of combining yaw control with axial induction control. To the best knowledge of the authors, the current manuscript presents a first attempt at quantifying the potential of combining the latter control strategies in a dynamic manner. 
The manuscript is structured as follows. Firstly, Section \ref{sec:meth} introduces the optimal control problem and the numerical methods used to solve it. Next, Section~\ref{sec:opt_yaw_setup} describes the simulation cases, and details setup parameters. Thereafter, Section~\ref{sec:opt_yaw_results} presents optimization results for both uniform and turbulent inflow, and identifies distinct yaw control regimes resulting in increased power extraction. Finally, Section~\ref{sec:opt_yaw_concl} summarizes the main findings of this manuscript and formulates suggestions for future research.
